---
title: "Analysis"
author: "Nico Bast"
date: "23 2 2021"
output: html_document

---

```{r setup, include=FALSE, cache=FALSE}

#options for all chunks
knitr::opts_chunk$set(echo = FALSE) #reflect code in resulting markdown
knitr::opts_chunk$set(cache = TRUE) #caching
knitr::opts_chunk$set(autodep = TRUE) #to remove error for caching in mardown result 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_knit$set(root.dir = "..") ##set root dir to project folder

#load packagaes
suppressMessages({
#preprocessing:
#require(openxlsx) #read excel files
require(readxl)  #read excel files
require(zoo) #required for na.approx func within pd.preprocess.func
require(plyr) #rbind.fill
require(papaja) #cite_r
  #analyses:
require(ggplot2)
require(lme4) #linear mixed model (LMM)
require(lmerTest) #LMM p-values
require(emmeans) #contrasts in LMM
require(hexbin) #see figure 3
  
#plotting
require(grid)
require(gridExtra)
require(kableExtra)
require(webshot) #save kable to file
  #webshot::install_phantomjs() #run manually to install phantomjs to save kable to file
})

#define paths
ifelse(Sys.info()['sysname']=='Linux',
       home_path<-'~',
       home_path<-'C:/Users/Nico')

folder_to_data<-"/PowerFolders/data_RING"
data_modality<-"/data_PDmemory"

path_to_data<-paste0(home_path,folder_to_data,data_modality)
path_to_project<-paste0(home_path,'/PowerFolders/Paper_RingMemoryPD')

path_to_data
path_to_project

```

# Load preprocessed data

```{r load_preprocessed_data}

load(paste0(path_to_project,'/data/main_df_preprocessed_220221.Rdata'))

ls()

```

# figure 3: preprocessed data

```{r figure_preprocessed_data, warning=FALSE, fig.height=4,fig.width=10}
#histogram of relative pupil dialtion (dependent variable)
d1<-ggplot(df,aes(x=rpd))+
      geom_histogram(fill='royalblue4',color='lightblue',bins=50)+ #GEOM
      xlim(0.8,1.3)+
      labs(x='relative pupil dilation (standardized)')+
      theme_bw()

#heatmap gaze behavior
g<-ggplot(df[df$ts.trial>30&df$ts.trial<=240 & (!is.na(df$pd)),],aes(x=XGazePosLeftEye,y=YGazePosLeftEye,fill=..density..))
d2<-g+geom_hex(bins=100)+xlim(0.4,0.6)+ylim(0.35,0.65)+scale_fill_gradientn(colours=rev(rainbow(3)))+
    labs(x='x-axis (standardized)',y='y-axis (standardized)')+
    theme_bw()


  #create file
  tiff(filename = paste0(path_to_project,"/output/figure3_histogram_heatmap.tiff"),
       width = 10, height = 4, units = "in", pointsize = 12, res=300, compression="lzw") #define dimensions


  grid.arrange(d1, d2, ncol=2)


  #save to file
  dev.off()
  
  grid.arrange(d1, d2, ncol=2)
  
```


# Sample description

```{r sample_description}


func_describe_table<-function(data,var){
  var_meanSD_ASD<-do.call(rbind,with(data,by(var,ASD,psych::describe)))['ASD',c('mean','sd')]
  var_range_ASD<-do.call(rbind,with(data,by(var,ASD,psych::describe)))['ASD',c('min','max')]
  var_range_ASD<-with(var_range_ASD,paste(round(min,2),'-',round(max,2)))  
  var_meanSD_TD<-do.call(rbind,with(data,by(var,ASD,psych::describe)))['TD',c('mean','sd')]
  var_range_TD<-do.call(rbind,with(data,by(var,ASD,psych::describe)))['TD',c('min','max')]
  var_range_TD<-with(var_range_TD,paste(round(min,2),'-',round(max,2)))  
  outcome<-cbind(round(var_meanSD_ASD,2),var_range_ASD,round(var_meanSD_TD,2),var_range_TD)
  row.names(outcome)<-deparse(substitute(var))
  outcome
}

df_describe<-rbind(
  func_describe_table(df_demo,df_demo$age),
  func_describe_table(df_demo,df_demo$VIQ),
  func_describe_table(df_demo,df_demo$PIQ),
  func_describe_table(df_demo,df_demo$FIQ),
  func_describe_table(df_demo,df_demo$AQ),
  func_describe_table(df_demo,df_demo$acc_agg),
  func_describe_table(df_demo,df_demo$bps_enc_agg),
  func_describe_table(df_demo,df_demo$bps_agg),
  func_describe_table(df_demo,df_demo$rpd_enc_agg),
  func_describe_table(df_demo,df_demo$rpd_rec_agg)
  )

row.names(df_describe)<-c('age','VIQ','PIQ','FIQ','AQ','recognition accuracy','BPS encoding','BPS retrieval','EEPR','SEPR')
names(df_describe)<-c('mean','SD','range','mean','SD','range')

func_group_difference<-function(data,var,group){
  anova_result<-with(data,summary(aov(var~group)))[[1]]
  t_value<-round(anova_result['F value'][1,1],2)
  p_value<-round(anova_result['Pr(>F)'][1,1],2)
  cohend_result<-round(with(data,psych::cohen.d(var,as.factor(group)))[[1]],2)
  cohend_d<-cohend_result[1,2]
  cohend_CI<-paste(cohend_result[1,1],',',cohend_result[1,3])
  cbind(t_value,p_value,cohend_d,cohend_CI)
}

df_group_difference<-data.frame(rbind(
  func_group_difference(df_demo,df_demo$age,df_demo$ASD),
  func_group_difference(df_demo,df_demo$VIQ,df_demo$ASD),
  func_group_difference(df_demo,df_demo$PIQ,df_demo$ASD),
  func_group_difference(df_demo,df_demo$FIQ,df_demo$ASD),
  func_group_difference(df_demo,df_demo$AQ,df_demo$ASD),
  func_group_difference(df_demo,df_demo$acc_agg,df_demo$ASD),
  func_group_difference(df_demo,df_demo$bps_enc_agg,df_demo$ASD),
  func_group_difference(df_demo,df_demo$bps_agg,df_demo$ASD),
  func_group_difference(df_demo,df_demo$rpd_enc_agg,df_demo$ASD),
  func_group_difference(df_demo,df_demo$rpd_rec_agg,df_demo$ASD)
  ))

names(df_group_difference)<-c('F','p','d','CI-95%')

df_describe<-cbind(df_describe,df_group_difference)

table1<-df_describe %>%
    kbl(caption = "Table 1. Sample description") %>%
    kable_classic(full_width = F, html_font = "Cambria") %>%
    add_header_above(c(" " = 1, "ASD (21m, 3f)" = 3, "TD (23m, 7f)" = 3, 'difference'=2, 'effect size' = 2))

table1

save_kable(table1,file=paste0(path_to_project,'/output/table1_sampledescription.pdf'))

```

# Recognition accuracy

```{r accuracy_between_groups}

 df_trial_old$PIC<-droplevels((df_trial_old$PIC))
  with(df_trial_old,round(table(group,Acc)/rowSums(table(group,Acc)),2))

    with(df_demo,by(acc_agg,ASD,psych::describe))
    t.test(acc_agg~ASD,df_demo)

  model_lmm<-glmer(Acc~group+
                    (1|word)+(1|PIC),family= 'binomial',data = df_trial_old)

  summary(model_lmm)
  exp(cbind(round(fixef(model_lmm)['groupTD'],2),
        round(confint(model_lmm,parm ='groupTD'),2)))

    confint(contrast(emmeans(model_lmm,~group),method='revpairwise'))
    ##--> higher accuracy in TD
  exp(fixef(model_lmm))

```

# EEPR between groups

## EEPR model summary

```{r EEPR}
#EEPR
anova(lmer(scale(rpd_enc)~group+(1|PIC)+(1|stimulus),data=df_trial_old))
summary(model_lm<-lmer(scale(rpd_enc)~group+(1|PIC)+(1|stimulus),data=df_trial_old))
confint(contrast(pmmeans(model_lm,~group),'pairwise')) #encoding pd effect

#BPS encoding
anova(lmer(scale(pd)~group+(1|PIC)+(1|stimulus),data=df_trial_enc))
summary(model_lm<-lmer(scale(pd)~group+(1|PIC)+(1|stimulus),data=df_trial_enc))



```

## figure 4: EEPR

```{r figure EEPR, warning=FALSE, fig.height=5, fig.width=9}

tiff(filename = paste0(path_to_project,"/output/figure4_encodingeffect_group.tiff"),
       width = 9, height = 5, units = "in", pointsize = 12, res=300, compression="lzw") #define dimensions

  g1<-ggplot(df_trial_old,aes(x=group,y=scale(rpd_enc),color=group))+
    geom_jitter(alpha=0.8)+geom_boxplot(fill='lightgray',notch = T,outlier.shape = NA)+
    ylim(-2,2)+
    xlab('')+ylab('EEPR (z)')+
    theme_bw()+
    theme(legend.position = "none")

  #Plotting
  #g<-ggplot(df_enc[df_enc$ts.trial<360,],aes(x=ts.trial*8.3/1000,y=rpd,color=group))
  g2<-ggplot(df_enc[df_enc$ts.trial>4 & df_enc$ts.trial<240,],aes(x=ts.trial*8.3/1000,y=scale(rpd),color=group))+
    geom_smooth(se=F, cex=2)+labs(x='seconds (s)',y='EEPR (z)')+
    theme_bw()+
    theme(legend.position = "none")+
    #geom_vline(xintercept = 2,lty=2)+
    stat_summary(fun.data = mean_se, geom = "errorbar")
  #facet_wrap(~stimulus)

  grid.arrange(g1,g2,widths= 1:2, ncol=2)

  dev.off()
  
  grid.arrange(g1,g2,widths= 1:2, ncol=2)

```

# SEPR between groups

## SEPR model summary

```{r SEPR}
#SEPR
anova(lmer(scale(rpd_response)~group*oldnew+(1|PIC)+(1|word),data=df_trial))
summary(model_lmm<-lmer(scale(rpd_response)~group*oldnew+(1|PIC)+(1|word),data=df_trial))
confint(contrast(pmmeans(model_lm,~group),'pairwise'))

cbind(round(fixef(model_lmm)['oldnewold'],2),
          round(confint(model_lmm,parm ='oldnewold'),2))

confint(contrast(emmeans(model_lmm,~oldnew|group),method='revpairwise'))

#BPS retrieval
anova(lmer(scale(pd)~group+(1|PIC)+(1|word),data=df_trial))
summary(model_lm<-lmer(scale(pd)~group+(1|PIC)+(1|word),data=df_trial))
```

## figure 5: SEPR 

```{r SEPR_figure, warning=FALSE, fig.height=5, fig.width=9}

 #create file
  tiff(filename = paste0(path_to_project,"/output/figure5_memoryeffect_group.tiff"),
       width = 9, height = 5, units = "in", pointsize = 12, res=300, compression="lzw") #define dimensions

  g1<-ggplot(df_trial_old,aes(x=group,y=scale(rpd_sepr_2s),color=group))+
    geom_jitter(alpha=0.8)+geom_boxplot(fill='lightgray',notch = T,outlier.shape = NA)+
    ylim(-2,2)+
    xlab('')+ylab('SEPR (z)')+
    theme_bw()+
    theme(legend.position = "none")

  #relative PD (divisive correction)
  g2<-ggplot(df[df$ts.trial<240&df$ts.trial>30,],aes(x=ts.trial*8.3/1000,y=scale(rpd),color=group,lty=oldnew))+
  #plot figure
  geom_smooth(se=F, cex=2)+
    stat_summary(fun.data = mean_se, geom = "errorbar", inherit.aes=F, aes(x=ts.trial*8.3/1000,y=scale(rpd),group=interaction(group,oldnew),color=group),cex=0.5)+ #inherit.aes=F and delivered aes as thios prevents different lty for errorbars
    labs(x='seconds (s)',y='SEPR (z)')+
    annotate('rect',ymin=0, ymax=0.75, xmin=1.75, xmax=2.00, alpha=0.2)+
    theme_bw()
    #facet_wrap(~factor(word,levels=c(1,2,3,4),labels=c('word','non-word','picture','shape')),ncol=2)

  grid.arrange(g1,g2,widths= 1:2, ncol=2)

  #save to file
  dev.off()

  grid.arrange(g1,g2,widths= 1:2, ncol=2)
  
```


# Effects on recognition accuracy

```{r effects_on_accuracy_logistic_regression}

###--> NOTE: we look at a logistic regression model
## - thus GLM logarithmizes binomial outcome variable to get a broader range (-inf, inf)
## thus all coefficients are log odds values, in order to translate these to probability, use the following formula

logit2prob <- function(logit){
  odds <- exp(logit)
  prob <- odds / (1 + odds)
  return(prob)
}

    #scaling
    df_trial_old$rpd_enc_scaled<-ifelse(scale(df_trial_old$rpd_enc)>3,NA,
                                        ifelse(scale(df_trial_old$rpd_enc)<(-3),NA,
                                               scale(df_trial_old$rpd_enc)))

    df_trial_old$rpd_response_scaled<-ifelse(scale(df_trial_old$rpd_response)>3,NA,
                                        ifelse(scale(df_trial_old$rpd_response)<(-3),NA,
                                               scale(df_trial_old$rpd_response)))


    ##logistic regression
    model_lr<-glmer(Acc~rpd_enc_scaled*rpd_response_scaled*group+
                      (1|PIC)+(1|stimulus),family= 'binomial',
                    data = df_trial_old)
    
 
    

    #plot table
      model_lr_summary<-data.frame(round(summary(model_lr)[['coefficients']],3)) ###--> as table
      row.names(model_lr_summary)<-c('intercept','EEPR (z)','SEPR (z)','group (TD)','EEPR x SEPR','EEPR x group','SEPR x group','EEPR x SEPR x group')
      colnames(model_lr_summary)<-c('estimate','std error','z-value','p-value')
      
      table2<-model_lr_summary %>%
          kbl(caption = "Table 2. Logistic regression: EEPR and SEPR effect on recognition accuracy") %>%
          kable_classic(full_width = F, html_font = "Cambria")
      
      
      save_kable(table2,file=paste0(path_to_project,'/output/table2_model_logistic_regression.pdf'))
      
      table2
      
         #effects
    emtrends(model_lr,~rpd_enc_scaled|group,var='rpd_response_scaled',at=list(rpd_enc_scaled = c(-2,-1,0,1,2)))
    confint(pmtrends(model_lr,~group,var='rpd_enc_scaled'))
    confint(pmtrends(model_lr,~group,var='rpd_response_scaled'))
      #--> no effect of SEPR if outliers are excluded
          
```

# figure6: three-way-interaction

```{r figure_effects_on_accuracy_posthoc, warning=FALSE, fig.height=4, fig.width=8}

      model_plot<-as.data.frame(emtrends(model_lr,~rpd_enc_scaled|group,var='rpd_response_scaled',at=list(rpd_enc_scaled = c(-2,-1,0,1,2))))
      g1<-ggplot(model_plot,aes(x=as.factor(rpd_enc_scaled)))+
        geom_hline(yintercept = 0)+
        geom_boxplot(aes(fill=group,
                         middle=rpd_response_scaled.trend,
                         lower=rpd_response_scaled.trend-SE,
                         upper=rpd_response_scaled.trend+SE,
                         ymin=asymp.LCL,
                         ymax=asymp.UCL),stat = "identity")+
        xlab('EEPR (z)')+ylab('SEPR effect (z)')+
        theme_bw()


tiff(filename = paste0(path_to_project,'/output/figure6_PDeffects_on_Acc.tiff'),
           width = 8, height = 4, units = "in", pointsize = 12, res=300, compression="lzw") #define dimensions

      g1
      dev.off()


      g1
      
    
      #model_lmm<-lmer(rpd_response_scaled~rpd_enc_scaled*group+
      #                  (1|PIC)+(1|stimulus),data = df_trial_old)
      #summary(model_lmm)

```


# remember and not remembered items

```{r remember_vs_notremember}

# remembered<-with(df_trial_old,ifelse(oldnew == 'old' & Acc == 0,'not_remembered',ifelse(oldnew =='old' & Acc == 1,'remembered',NA)))
# 
# lm_model<-glmer(as.factor(remembered)~rpd_enc_scaled*group+rpd_response_scaled*group+(1|PIC)+(1|word),family= 'binomial',
#                     data = df_trial_old)
# 
# summary(lm_model)
# emtrends(lm_model,~group,var='rpd_response')
#     
# 
# table(is.na(remembered))

```


# supplements: separate models on accuracy

```{r effects_on_accuracy_separate_models}

    ###-> split into two models
    #ASD model:
    model_lr<-glmer(Acc~scale(rpd_enc)+scale(rpd_response)+
                      (1|PIC)+(1|stimulus),family= 'binomial',
                    data = df_trial_old[df_trial_old$group=='ASD',])

     model_lr_summary<-data.frame(round(summary(model_lr)[['coefficients']],3)) ###--> as table
     row.names(model_lr_summary)<-c('intercept','EEPR','SEPR')
     
      table_supplements<-model_lr_summary %>%
        kbl(caption = "logistic regression in ASD: EEPR and SEPR effect on recognition accuracy") %>%
        kable_classic(full_width = F, html_font = "Cambria")

      save_kable(table_supplements,file=paste0(path_to_project,'/output/supplements_table3.pdf'))
      table_supplements
      
    #--> NOTE: LR coefficiens are LOG ODDS : log(acc/(1-acc)) --> convert to likelihood: exp(x)/(1+exp())
    logit2prob(fixef(model_lr))
    exp(fixef(model_lr))
    confint(model_lr,parm='scale(rpd_enc)')
    confint(model_lr,parm='scale(rpd_response)')
  
    #TD model
    model_lr<-glmer(Acc~scale(rpd_enc)+scale(rpd_response)+
                      (1|PIC)+(1|stimulus),family= 'binomial',
                    data = df_trial_old[df_trial_old$group=='TD',])
    
    model_lr_summary<-data.frame(round(summary(model_lr)[['coefficients']],3)) ###--> as table
    row.names(model_lr_summary)<-c('intercept','EEPR','SEPR')
    
    
    table_supplements<-model_lr_summary %>%
      kbl(caption = "logistic regression in TD: EEPR and SEPR effect on recognition accuracy") %>%
      kable_classic(full_width = F, html_font = "Cambria")
    
    save_kable(table_supplements,file=paste0(path_to_project,'/output/supplements_table4.pdf'))
      table_supplements

    #--> NOTE: LR coefficiens are LOG ODDS : log(acc/(1-acc)) --> convert to likelihood: exp(x)/(1+exp())
    logit2prob(fixef(model_lr))
    exp(fixef(model_lr))
    confint(model_lr,parm='scale(rpd_enc)')
    confint(model_lr,parm='scale(rpd_response)')

```

# supplements: data distribution

```{r distribution_PD}

theme_set(theme_bw())

grid.arrange(
  ggplot(df_trial,aes(bps_enc_agg))+geom_histogram(bins=10)+xlab('BPS encoding'),
  ggplot(df_trial,aes(bps_agg))+geom_histogram(bins=10)+xlab('BPS retrieval'),
  ggplot(df_trial_old,aes(rpd_response))+geom_histogram(bins=20)+xlab('EEPR'),
  ggplot(df_trial,aes(rpd_response))+geom_histogram(bins=20)+xlab('SEPR'),
  nrow=2
)

```


